---
title: "Week6_HW_PL"
author: "Peter Lombardo"
date: "March 9, 2019"
output: html_document
---
```{r}
suppressMessages(suppressWarnings(library(readxl)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(forecast)))
suppressMessages(suppressWarnings(library(fpp2)))
suppressMessages(suppressWarnings(library(mlbench)))
suppressMessages(suppressWarnings(library(caret)))
suppressMessages(suppressWarnings(library(caTools)))
```

#8.1

##A
Explain the differences among these figures. Do they all indicate that the data are white noise?

Yes because because there is no correlation from one point to the next. 

##B
Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?

It can be assumed that some observations can break the critical values while stil being white noise. 

#8.2
A classic example of a non-stationary series is the daily closing IBM stock price series (data set ibmclose). Use R to plot the daily closing prices for IBM stock and the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.
```{r}
data(ibmclose)
tsdisplay(ibmclose)
```
the ACF plot shows strong autocorrelation.

#8.7
Consider wmurders, the number of women murdered each year (per 100,000 standard population) in the United States.
```{r}
data(wmurders)
```

##A
```{r}
tsdisplay(wmurders)
```
##B
Should you include a constant in the model? Explain.

No, there is no drastic chnage in the avergae. 

##C
Write this model in terms of the backshift operator.

(1-B-B2)(1-B)yt

##D
Fit the model using R and examine the residuals. Is the model satisfactory?
```{r}
fit <- Arima(wmurders, order = c(2,1,0))
plot(fit$residuals)
```
There are no pattern in the resdiual, therefore the modle is satisfactory.

##E
Forecast three times ahead. Check your forecasts by hand to make sure that you know how they have been calculated.
```{r}
forecast <- forecast(fit, h=3)
forecast
```
##F
```{r}
plot(forecast)
```
##G
```{r}
fit_g <- auto.arima(wmurders)
fit_g
```

The modles appear to be different. I would assume this modle is better since it is auto calculated. 

#8.12
```{r}
data(mcopper)
plot(mcopper)
```
##A
```{r}
lamda <- BoxCox.lambda(mcopper)
mcopper_bcx <- BoxCox(mcopper, lambda = lamda)
tsdisplay(mcopper_bcx)
```
##B
```{r}
fit <- auto.arima(mcopper, trace = TRUE, ic ="aic", lambda = lamda)
fit
```
##C
```{r}
fit_2 <- Arima(mcopper, order = c(2,1,2), lambda = lamda)
fit_3 <- Arima(mcopper, order = c(0,1,1), lambda = lamda)
accuracy(fit)
accuracy(fit_2)
accuracy(fit_3)
```
##D
```{r}
plot(fit_2$residuals)
```
##E
```{r}
forecast <- forecast(fit_2, h=10)
plot(forecast)
```
The forecast look reasonable. 
##F
```{r}
fit_4 <- ets(mcopper); fit_4
forecast2 <- forecast(fit_4, h=10)
plot(forecast2)
```