---
title: '7.5'
author: "Alejandro D. Osborne"
date: "April 24, 2019"
output: html_document
---
```{r}
library(AppliedPredictiveModeling)
library(caret)
data(ChemicalManufacturingProcess)
```

(a)
```{r}
predictors <- subset(ChemicalManufacturingProcess,select= -Yield)
yield <- subset(ChemicalManufacturingProcess,select="Yield")
```

```{r}
set.seed(517)
trainingRows <- createDataPartition(yield$Yield,
p = 0.7,
list = FALSE)
```

```{r}
trainPredictors <- predictors[trainingRows,]
trainYield <- yield[trainingRows,]
```

```{r}
testPredictors <- predictors[-trainingRows,]
testYield <- yield[-trainingRows,]
```

```{r}
#Pre-process trainPredictors and apply to trainPredictors and testPredictors
pp <- preProcess(trainPredictors,method=c("BoxCox","center","scale","knnImpute"))
ppTrainPredictors <- predict(pp,trainPredictors)
ppTestPredictors <- predict(pp,testPredictors)
```

```{r}
#Identify and remove NZV
nzvpp = nearZeroVar(ppTrainPredictors)
ppTrainPredictors <- ppTrainPredictors[-nzvpp]
ppTestPredictors <- ppTestPredictors[-nzvpp]
```

```{r}
#Identify and remove highly correlated predictors
predcorr = cor(ppTrainPredictors)
```

```{r}
highCorrpp <- findCorrelation(predcorr)
ppTrainPredictors <- ppTrainPredictors[, -highCorrpp]
ppTestPredictors <- ppTestPredictors[, -highCorrpp]
```

```{r}
#Set-up trainControl
set.seed(517)
ctrl <- trainControl(method = "boot", number = 25)
```

```{r}
set.seed(614)

marsChemGrid <- expand.grid(degree = c(1:2), nprune = c(2:10))
marsChemTune <- train(x = ppTrainPredictors, y = trainYield,
method = "earth",
trControl = ctrl,
tuneGrid = marsChemGrid)

psvmTuneGrid <- expand.grid(C=c(0.01,0.05,0.1), degree=c(1,2), scale=c(0.25,0.5,1))
PSVMChemTune <- train(x = ppTrainPredictors, y = trainYield,
method = "svmPoly",
trControl = ctrl,
tuneGrid = psvmTuneGrid)
```


(b)

```{r}
marsChemTune$results$degree[best(marsChemTune$results, "Rsquared", maximize = TRUE)]
```

```{r}
marsChemTune$results$nprune[best(marsChemTune$results, "Rsquared", maximize = TRUE)]
```

```{r}
round(marsChemTune$results$Rsquared[best(marsChemTune$results, "Rsquared", maximize = TRUE)],2)
```

##The optimal degree and number of terms that maximize R2 are 1 and 3, respectively, with an R2 of 0.51
##The top two PLS predictors are ManufacturingProcess32 and ManufacturingProcess09, which are the same as what the MARS
model identifies.

```{r}
chemMARSSummary <- summary(marsChemTune$finalModel)
chemMARSSummary
```


```{r}
varImp(marsChemTune)
```

The final MARS model has slightly worse cross-validation and test set performance than the optimal PLS model. This indicates the underlying structure between the predictors and the response is approximately linear.  


 the top two PLS predictors are ManufacturingProcess32 and ManufacturingProcess09, which are the same as what the MARS model identifies.  PLS, however, identifies additional predictive information from the other predictors that improve the predictive ability of the models.  Overall, many of the manufacturing process predictors are at the top of the importance list.

indicates that the optimal degree, cost, and scale that maximize $R^2$ are 

```{r}
PSVMChemTune$results$degree[best(PSVMChemTune$results, "Rsquared", maximize = TRUE)]
```
```{r}
PSVMChemTune$results$C[best(PSVMChemTune$results, "Rsquared", maximize = TRUE)]
```
```{r}
PSVMChemTune$results$scale[best(PSVMChemTune$results, "Rsquared", maximize = TRUE)]
```
respectively, with an $R^2$ of 

```{r}
round(PSVMChemTune$results$Rsquared[best(PSVMChemTune$results, "Rsquared", maximize = TRUE)],2)
```

```{r}
plotTheme <- bookTheme()
trellis.par.set(plotTheme)
plot(PSVMChemTune,metric="Rsquared")
```

The polynomial SVM tuning parameters both indicate that a linear model is sufficient for this model.  Specifically, a degree of 1 is an indicator of linear structure, and a low cost is also an indicator of linear structure.

