---
title: "HW3"
output: html_document
---

```{r}
library(mlbench)
library(e1071)
library(corrplot)
library(caret)
```
## 3.1
```{r}
data(Glass)
str(Glass)
```

```{r}
names<-names(Glass)
classes<-sapply(Glass,class)

for(name in names[classes == 'numeric'])
{

    hist(Glass[,name],xlab =name , main =paste("Hstogram of ",  name) ) 
}

```
```{r}
Glass_F <-Glass[,-which(names(Glass)=="Type")]
skew<-apply(Glass_F,2,skewness)
skew

```
```{r}
correlations<-cor(Glass_F)
corrplot(correlations,order="hclust")
correlations
```
```{r}
for (name in names(Glass_F))
boxplot(Glass_F[,name],xlab = name)

```
from these plot above I think some pridectors such as K and Fe may have vaild outliers. in addtion predicor RI and Ca appear to have high postive correlation and some predictors have negative correlation like RI and SI 

in terms of skewness, none predictor are prefectly symmetric. predictor K,Ba is highly right skewed and Ca is also right skewed. among all predictors, AI seems to be relative symmetric.

```{r}
Glass_AT <- Glass_F
Glass_T<-apply(Glass_F,2,BoxCoxTrans)
i <-1
for (name in names(Glass_F))
{
Glass_AT[,name] <-predict(Glass_T[i],Glass_F[,name]) 
i <- i+1
  
}
skew_AT<-apply(Glass_AT,2,skewness)
skew_AT
```
```{r}
for(name in names[classes == 'numeric'])
{

    hist(Glass_AT[,name],xlab =name , main =paste("Hstogram of ",  name) ) 
}
```

Due to 0 value some predictors are not transformed such as K,Ba,Fe and for those transformed predictor the skewness are improved a little bit the predictor benefit most from the transformation is Ca


---
title: "Week3_HW_PL"
output: html_document
---
#3.1
```{r}
suppressMessages(suppressWarnings(library(mlbench)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(tidyr)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(ggplot2)))
```
```{r}
data(Glass)
str(Glass)
```
##a
```{r}
Glass
```
```{r}
ggplot(Glass,aes(x=Type))+
  geom_bar(colour="black")+
  ggtitle("Observation Count by Glass Type")
```
```{r}
for(i in 1:9){
  x <- Glass[,i]
   hist(x,
       main=paste("Series Chart",i,": Box Plot of Chemical Elements",
          names(Glass)[i],"by Type of Glass"),
       xlab="Type of Glass",
       col = ("Dark Green"))
}
```
#3.2
```{r}
data(Soybean)
```

```{r}
Soybean
```
##A
```{r}
na_count <-sapply(Soybean, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na_count$names <- rownames(na_count)
na_count
```
```{r}
ggplot(data=na_count) + geom_point(aes(x=na_count,y=names))
```
##B
```{r}
#Seed, Fruit, and Leef data seems to be the most scarce
```
##C
```{r}
Soybean_NA_Fix<- Soybean[is.na(Soybean)] <- 0
Soybean_NA_Fix
na_count_2 <-sapply(Soybean, function(y) sum(length(which(is.na(y)))))
na_count_2 <- data.frame(na_count_2)
na_count_2$names <- rownames(na_count_2)
na_count_2
```