---
title: "HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(fpp2)
```

## 7.5
```{r}
autoplot(books) + ggtitle("Daily Sales") + xlab("Day") 
hard <- books[,"Hardcover"]
fch <- ses(hard, h=5)
round(accuracy(fch),2)
autoplot(fch) +
  autolayer(fitted(fch), series="Fitted") + ggtitle("Daily Sales for Hardcover Books") +
  ylab("books") + xlab("Year")
paper <- books[,"Paperback"]
fcp <- ses(paper, h=5)
round(accuracy(fcp),2)
autoplot(fcp) +
  autolayer(fitted(fcp), series="Fitted") + ggtitle("Daily Sales for Papercover Books") +
  ylab("books") + xlab("Year")
```
Both series has a upward trend and seasonality and the variance seasonality changes overtime 

## 7.6
```{r}
fchh <- holt(hard, h=5)
autoplot(hard) +
  autolayer(fchh, series="Holt's method", PI=FALSE) +
  ggtitle("Forecasts for Hard Books from Holt's method") + xlab("Day") +
  ylab("Books") +
  guides(colour=guide_legend(title="Forecast"))
round(accuracy(fchh),2)
fcph <- holt(paper, h=5)
autoplot(paper) +
  autolayer(fcph, series="Holt's method", PI=FALSE) +
  ggtitle("Forecasts for Paper Books from Holt's method") + xlab("Day") +
  ylab("Books") +
  guides(colour=guide_legend(title="Forecast"))
round(accuracy(fcph),2)
```
according to the RMSE maesure, the Holt Linear method work better for both series. BUt from the forcast plot, I think the Holt Linear method is better than simple exponential smoothing forpaper book becasue it realized the increase trend for the series and I think simple exponential smoothing works better for hard cover becasue it interpret the fluctuation 

```{r}
fcp$Lo <- fcp$fitted - 1.96*33.64
fcp$Hi <- fcp$fitted + 1.96*33.64
fcph$Lo <-fcph$fitted - 1.96*31.14 
fcph$Hi <-fcph$fitted + 1.96*31.14 
fch$Lo <- fch$fitted - 1.96*31.93
fch$Hi <- fch$fitted + 1.96*31.93
fchh$Lo <- fchh$fitted - 1.96*27.19
fchh$Hi <- fchh$fitted + 1.96*27.19

cbind(fcp$Lo,fcp$Hi,fcph$Lo,fcph$Hi)
cbind(fch$Lo,fch$Hi,fchh$Lo,fchh$Hi)
```
the Holt's method has a smaller interval since it has a lower RMSE

## 7.10
##A B
```{r}
autoplot(ukcars) + ggtitle("UK passenger vehicle production") + xlab("year")

ukcars  %>%
  stl(t.window=4, s.window="periodic", robust=TRUE) %>%
  autoplot() + ggtitle("UK passenger vehicle production")
ukcars  %>%
  stl(t.window=4, s.window="periodic", robust=TRUE)->fit
autoplot(ukcars, series="Data") +
autolayer(seasadj(fit), series="Seasonally Adjusted") + xlab("Year") + ggtitle("Seasonally Adjusted UK passenger vehicle production")
```
the series has a obvious trand and seasonality and the variance of seasonality is not constant

## C
```{r}
stlf(seasadj(fit),etsmodel="AAN", damped=TRUE) %>% autoplot() 

```

##D
```{r}
holt(seasadj(fit), h=8, damped = FALSE)%>% autoplot() 
```
##E
```{r}
fit2<-ets(seasadj(fit))
fit2 %>% forecast(h=8) %>%
  autoplot() 
```
##F
```{r}
fit1<-stlf(seasadj(fit),etsmodel="AAN", damped=TRUE) 
summary(fit2)
summary(fit1)
```
seems like stl model is better since RMSE is lower

```{r}
fit3<-holt(seasadj(fit), h=8, damped = FALSE)
summary(fit3)
```
compare all three approch, the stl model looks the best
```{r}
checkresiduals(fit1)
checkresiduals(fit2)
checkresiduals(fit3)
```
from the residual analysis, seems like stl model still have some correlations but is indeed the best among all three models
